---
title: "DMD mdx corticosteroids project: Statistical analysis - Compare genotypes at T0"
author: Simina Boca
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "figures/")
```

```{r chunk1}
library(dplyr)
library(reshape2)
library(ggplot2)
library(nlme)
library(qvalue)
source("functions.R")
```

## Read in data

```{r chunk2}
MasterFrame <- read.delim(file="../data/Supplemental Table 1.tsv",
                          header=TRUE,
                          sep="\t",
                          check.names=FALSE)
```

## Prepare data 

Prepare data for analysis:

### Take out QC samples

```{r chunk3}
##first take out the QC samples
MasterFrameNoQC <- MasterFrame[MasterFrame$SampleID != "QC",]
##only include interaction terms between Times T1/T2 and Group variables (since just one group 2 at baseline)
MasterFrameNoQC$TimeT1 <- MasterFrameNoQC$TimeT2 <- 
  MasterFrameNoQC$GenoMDX <- 
  MasterFrameNoQC$GroupPred <- 
  rep(0, nrow(MasterFrameNoQC)) 
MasterFrameNoQC$TimeT1[MasterFrameNoQC$Time=="T1"] <- 1
MasterFrameNoQC$TimeT2[MasterFrameNoQC$Time=="T2"] <- 1
MasterFrameNoQC$GenoMDX[MasterFrameNoQC$Genotype=="mdx23"] <- 1
MasterFrameNoQC$GroupPred[MasterFrameNoQC$Group=="group 1"] <- 1

dim(MasterFrame)
dim(MasterFrameNoQC)
```

### Remove samples with 0 variance at T0

```{r}

##get metabolites
mets <- colnames(MasterFrameNoQC)[grep("m/z",colnames(MasterFrameNoQC))]

##actually look at top 75% most variable metabolites
##metsMostVar <- names(sort(sapply(MasterFrameNoQC[,mets],var))[round(0.25*length(mets)):length(mets)])

##mets <- metsMostVar

MasterFrame_T0 <- subset(MasterFrameNoQC, Time=="T0")
##remove any metabolites with variance = 0
var0 <- names(which(apply(MasterFrame_T0[,mets],2,var)==0))
var0
##look at this peak!
unique(MasterFrame_T0[,var0])

MasterFrame_T0 <- select(MasterFrame_T0, -one_of(var0))
mets <- setdiff(mets, var0)

length(mets)
```

## Run t-tests

Run t-tests with assumption of equal variance:

```{r chunk4_equal_var}

t_tests_equal <- lapply(MasterFrame_T0[,mets],
                        function(metInt, geno){
                          ##full model
                          t.test(metInt[geno==1],metInt[geno==0],
                                 var.equal=TRUE)
                        }, 
                        MasterFrame_T0$GenoMDX)
```

Run t-tests with assumption of unequal variance:

```{r chunk4_diff_var}

t_tests_diff <- lapply(MasterFrame_T0[,mets],
                        function(metInt, geno){
                          ##full model
                          t.test(metInt[geno==1],metInt[geno==0],
                                 var.equal=FALSE)
                        }, 
                        MasterFrame_T0$GenoMDX)

```

Get p-values from both types of test to compare them:
```{r chunk5}
p_vals_equal <- sapply(t_tests_equal, function(t){t$p.value})
p_vals_diff <- sapply(t_tests_diff, function(t){t$p.value})

cor(p_vals_equal, p_vals_diff)
plot(p_vals_equal, p_vals_diff)

p_vals <- p_vals_equal
```

Take a big-picture view of these p-values:
```{r chunk6}
hist(p_vals, col="grey")
qvalue(p_vals)$pi0

round((1-qvalue(p_vals)$pi0)*100)

sum(p_vals < 0.05/length(p_vals))
```

## Look at creatine, creatinine, validated MS/MS peaks

Take a look at putative creatine and creatinine p-values:
```{r chunk7}
mets_114 <- names(p_vals)[grep("m/z = 114\\.(0|1)",names(p_vals))]
mets_114 <- mets_114[grep("Positive",mets_114)]
p_vals[mets_114]

mets_132 <- names(p_vals)[grep("m/z = 132\\.(0|1)",names(p_vals))]
mets_132 <- mets_132[grep("Positive",mets_132)]
p_vals[mets_132]
```

Make some plots for these metabolites plus the two ones that were validated in MS/MS:
```{r chunk8}
met_top <- names(sort(p_vals))[1]
met_top
met_creatinine <- mets_114
met_creatinine
met_creatine <- mets_132[grep("132.077",mets_132)]
met_creatine
met_hypusine <- "m/z = 234.18087, mode = Positive, RT = 19.39639"
met_biliverdin <- "m/z = 583.25604, mode = Positive, RT = 381.06653"
met_484 <- "m/z = 484.16091, mode = Negative, RT = 155.62501"
met_885 <- "m/z = 884.92954, mode = Positive, RT = 231.26267"

met_229 <- "m/z = 229.15522, mode = Positive, RT = 25.16714"
```
Change genotype from "mdx23" to "mdx":
```{r}
MasterFrame_T0$Genotype <- as.character(MasterFrame_T0$Genotype)
MasterFrame_T0$Genotype[MasterFrame_T0$Genotype == "mdx23"] <- "mdx"
```

```{r boxplots_geno_T0, fig.height=3, fig.width=16}
box1 <- ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_top,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  ggtitle("a)\nm/z = 357.25, Positive mode") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

box2 <- ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_creatinine,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  ggtitle("b)\nCreatinine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

box3 <- ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_creatine,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  ggtitle("c)\nCreatine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

box4 <- ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_hypusine,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  ggtitle("d)\nHypusine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

box5 <- ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_biliverdin,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  ggtitle("e)\nBiliverdin") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))

multiplot(box1, box2, box3, box4, box5, cols=5)
```

```{r boxplots_geno_T0_Supp_2}
ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_484,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```

```{r boxplots_geno_T0_Supp_3}
ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_885,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```

Look at metabolite 229:
```{r boxplots_met_229}
ggplot(MasterFrame_T0, aes_string(x="Genotype", y=paste("`",met_229,"`",sep=""))) +
  geom_boxplot(fill="grey") +
  geom_point() +
  ylab("Normalized, log2-transformed value") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```

Save everything:
```{r}
save(list=ls(), file = "../objs/stat_analysis_geno_T0.RData")
```

Dump top 50 peaks in a file:
```{r}
##see what rank creatine is at
rank(p_vals)[met_creatine]
p_vals[met_creatine]
p_vals[met_creatine]*length(p_vals)

p_vals <- sort(p_vals)
top_peaks <- names(which(rank(p_vals)<=50))
p_vals_top_peaks <- p_vals[top_peaks]

info_top_peaks <- strsplit(top_peaks, ", ")
mz_top_peaks <- as.numeric(gsub("m/z = ", "", sapply(info_top_peaks, function(x){x[1]})))
mode_top_peaks <- gsub("mode = ", "", sapply(info_top_peaks, function(x){x[2]}))
rt_top_peaks <- as.numeric(gsub("RT = ", "", sapply(info_top_peaks, function(x){x[3]})))

##also get difference in means
log2_fold_change <- sapply(t_tests_equal, function(t){means <- t$estimate; means[1]-means[2]})
names(log2_fold_change) <- names(t_tests_equal)

top_peaks_df <- data.frame(mz = mz_top_peaks, mode = mode_top_peaks, 
                           rt = rt_top_peaks,
                           p_value = p_vals_top_peaks,
                           log2_fold_change = log2_fold_change[top_peaks])
##sort by mode (to make look-up easier)
top_peaks_df <- top_peaks_df[order(top_peaks_df$mode),]
write.csv(top_peaks_df, file="../top_peaks_compare_genotypes_T0_urine.csv",
          row.names = FALSE)
```

## Session info

```{r}
sessionInfo()
```
