---
title: "DMD mdx corticosteroids project: Statistical analysis - Compare top peaks from genotype comparison at T0 at T1-T2"
author: Simina Boca
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "figures/")
```

```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(nlme)
library(qvalue)
library(limma) #for vennDiagram function
source("functions.R")

```

## Load data, previous analysis results

Read in data, load results from genotype comparison at time T0:

```{r}
##load analysis results
load("../objs/stat_analysis_geno_T0.RData")

##clarify what the p-values are from
p_vals_T0 <- p_vals
##read in formatted data
MasterFrame <- read.delim(file="../data/Supplemental Table 1.tsv",
                          header=TRUE,
                          sep="\t",
                          check.names=FALSE)
```

## Prepare data 

Prepare data for analysis:

```{r}
##first take out the QC samples
MasterFrameNoQC <- MasterFrame[MasterFrame$SampleID != "QC",]
##only include interaction terms between Times T1/T2 and Group variables (since just one group 2 at baseline)
MasterFrameNoQC$TimeT1 <- MasterFrameNoQC$TimeT2 <- 
  MasterFrameNoQC$GenoMDX <- 
  MasterFrameNoQC$GroupPred <- 
  rep(0, nrow(MasterFrameNoQC)) 
MasterFrameNoQC$TimeT1[MasterFrameNoQC$Time=="T1"] <- 1
MasterFrameNoQC$TimeT2[MasterFrameNoQC$Time=="T2"] <- 1
MasterFrameNoQC$GenoMDX[MasterFrameNoQC$Genotype=="mdx23"] <- 1
MasterFrameNoQC$GroupPred[MasterFrameNoQC$Group=="group 1"] <- 1

##get metabolites
mets <- colnames(MasterFrameNoQC)[grep("m/z",colnames(MasterFrameNoQC))]

MasterFrame_T0 <- subset(MasterFrameNoQC, Time=="T0")
##remove any metabolites with variance = 0
var0 <- names(which(apply(MasterFrame_T0[mets],2,var)==0))
var0
MasterFrame_T0 <- dplyr::select(MasterFrame_T0, -one_of(var0))
mets <- setdiff(mets, var0)

length(mets)

##now only keep these metabolites for the T1 vs T2 comparison ie remove the variance 0 ones
MasterFrameNoQC <- dplyr::select(MasterFrameNoQC, -one_of(var0))
##only keep times T1 and T2
MasterFrame_T1_T2 <- subset(MasterFrameNoQC, Time!="T0")
MasterFrame_T1_T2$Time <- as.character(MasterFrame_T1_T2$Time)

all_mets <- mets
```

## Get some descriptive tables

Time T0:
```{r}
table(MasterFrame_T0$Genotype)
```

Time T1:
```{r}
table(MasterFrame_T1_T2$Group[MasterFrame_T1_T2$Time=="T1"],
      MasterFrame_T1_T2$Genotype[MasterFrame_T1_T2$Time=="T1"])
```

Time T2:
```{r}
table(MasterFrame_T1_T2$Group[MasterFrame_T1_T2$Time=="T2"],
      MasterFrame_T1_T2$Genotype[MasterFrame_T1_T2$Time=="T2"])
```

Overlap between samples at Time 1 and Time 2:
```{r}
samples_times <- table(MasterFrame_T1_T2$SampleID,
                       MasterFrame_T1_T2$Time)
length(unique(MasterFrame_T1_T2$SampleID))
colSums(samples_times)
sum(samples_times[,1]*(1-samples_times[,2]))
sum(samples_times[,2]*(1-samples_times[,1]))
```

## Make some plots for some of the known interesting metabolites
```{r}
met_top <- "m/z = 357.24594, mode = Positive, RT = 20.659"
met_creatine <- "m/z = 132.07757, mode = Positive, RT = 23.18143"
met_creatinine <- "m/z = 114.06447, mode = Positive, RT = 22.59731"
met_hypusine <- "m/z = 234.18087, mode = Positive, RT = 19.39639"
met_biliverdin <- "m/z = 583.25604, mode = Positive, RT = 381.06653"

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_top,"`",sep=""),
                                   group="SampleID", color="Group", linetype="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4))

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_top,"`",sep=""),
                                   group="SampleID", color="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  facet_grid(. ~ Group) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4)) 

ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_top,"`",sep=""),
                                     group="SampleID", color="Genotype", linetype="Group")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx23"),
                     labels = c("WT", "mdx23"),
                     values = c(2, 4))

############################################################

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_creatine,"`",sep=""),
                                   group="SampleID", color="Group", linetype="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4))

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_creatine,"`",sep=""),
                                   group="SampleID", color="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  facet_grid(. ~ Group) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4))

ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_creatine,"`",sep=""),
                                     group="SampleID", color="Genotype", linetype="Group")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx23"),
                     labels = c("WT", "mdx23"),
                     values = c(2, 4))

############################################################

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_creatinine,"`",sep=""),
                                   group="SampleID", color="Group", linetype="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4))

ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met_creatinine,"`",sep=""),
                                   group="SampleID", color="Group")) +
  geom_point() +
  geom_line(aes(color=Genotype)) +
  facet_grid(. ~ Group) +
  scale_color_manual(name = "Genotype and group", 
                     breaks = c("group 1", "group 2", "WT", "mdx23"),
                     labels = c("group 1", "group 2", "WT", "mdx23"),
                     values = c(1, 3, 2, 4))

ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_creatinine,"`",sep=""),
                                     group="SampleID", color="Genotype", linetype="Group")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx23"),
                     labels = c("WT", "mdx23"),
                     values = c(2, 4))
```

## Make the plots for the paper
Change genotype from "mdx23" to "mdx":
```{r}
MasterFrame_T1_T2$Genotype <- as.character(MasterFrame_T1_T2$Genotype)
MasterFrame_T1_T2$Genotype[MasterFrame_T1_T2$Genotype == "mdx23"] <- "mdx"

MasterFrame_T1_T2$Treatment <- MasterFrame_T1_T2$Group
MasterFrame_T1_T2$Treatment <- as.character(MasterFrame_T1_T2$Treatment)
MasterFrame_T1_T2$Treatment[MasterFrame_T1_T2$Treatment == "group 1"] <- "prednisolone"
MasterFrame_T1_T2$Treatment[MasterFrame_T1_T2$Treatment == "group 2"] <- "vehicle control"


```

```{r plots_T1_T2, fig.height=3, fig.width=25}

plot1 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_top,"`",sep=""),
                                              group="SampleID", color="Genotype", shape="Genotype", 
                                              linetype="Treatment")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(2, 4)) +
  scale_shape_manual(name="Genotype",
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(17, 16)) +
  ylab("Normalized, log2-transformed value") + 
  ggtitle("a)\nm/z = 357.25, Positive mode") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5),
        legend.text = element_text(size=11))

plot2 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_creatinine,"`",sep=""),
                                              group="SampleID", color="Genotype", shape="Genotype", 
                                              linetype="Treatment")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(2, 4)) +
  scale_shape_manual(name="Genotype",
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(17, 16)) +
  ylab("Normalized, log2-transformed value") + 
  ggtitle("b)\nCreatinine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5),
        legend.text = element_text(size=11))

plot3 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_creatine,"`",sep=""),
                                              group="SampleID", color="Genotype", shape="Genotype", 
                                              linetype="Treatment")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(2, 4)) +
  scale_shape_manual(name="Genotype",
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(17, 16)) +
  ylab("Normalized, log2-transformed value") + 
  ggtitle("c)\nCreatine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5),
        legend.text = element_text(size=11))

plot4 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_hypusine,"`",sep=""),
                                              group="SampleID", color="Genotype", shape="Genotype", 
                                              linetype="Treatment")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(2, 4)) +
  scale_shape_manual(name="Genotype",
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(17, 16)) +
  ylab("Normalized, log2-transformed value") + 
  ggtitle("d)\nHypusine") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5),
        legend.text = element_text(size=11))

plot5 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met_biliverdin,"`",sep=""),
                                              group="SampleID", color="Genotype", shape="Genotype", 
                                              linetype="Treatment")) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Genotype", 
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(2, 4)) +
  scale_shape_manual(name="Genotype",
                     breaks = c("WT", "mdx"),
                     labels = c("WT", "mdx"),
                     values = c(17, 16)) +
  ylab("Normalized, log2-transformed value") + 
  ggtitle("e)\nBiliverdin") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5),
        legend.text = element_text(size=11))

multiplot(plot1, plot2, plot3, plot4, plot5, cols=5)

```

Compare values of creatine at T0 for WT for Group 1 vs Group 2:

```{r}
MasterFrame_T0_WT <- filter(MasterFrame_T0, GenoMDX==0)
ggplot(MasterFrame_T0_WT, aes_string(x="Group",y=paste("`",met_creatine,"`",sep=""))) +
  geom_boxplot() + geom_point()
```

Are they significantly different from each other?

```{r}
t.test(MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 1",met_creatine],
       MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 2",met_creatine])

```

What about for the metabolite with weight of 357?

```{r}
ggplot(MasterFrame_T0_WT, aes_string(x="Group",y=paste("`",met_top,"`",sep=""))) +
  geom_boxplot() + geom_point()
```

Are they significantly different from each other?

```{r}
t.test(MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 1",met_top],
       MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 2",met_top])

```

## Test hypotheses relating to impact of genotype, group, and time for these metabolites

Fit models:

```{r}
mets <- c(met_top, met_creatine)
##full models
lme_full <- lapply(MasterFrame_T1_T2[,mets],
                    function(metInt, geno, group, Times, ID){
                      ##full model
                      lme(metInt ~ geno*Times*group, random =~ 1 | ID,
                          method="ML");
                    }, 
                    MasterFrame_T1_T2$GenoMDX, MasterFrame_T1_T2$GroupPred, 
                    MasterFrame_T1_T2$Time,
                    MasterFrame_T1_T2$SampleID)
##models with no genotypes
lme_no_geno <- lapply(MasterFrame_T1_T2[,mets],
                      function(metInt, group, Times, ID){
                        ##full model
                        lme(metInt ~ Times*group, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GroupPred, 
                      MasterFrame_T1_T2$Time,
                      MasterFrame_T1_T2$SampleID)
##models with no group
lme_no_group <- lapply(MasterFrame_T1_T2[,mets],
                      function(metInt, geno, Times, ID){
                        ##full model
                        lme(metInt ~ Times*geno, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GenoMDX, 
                      MasterFrame_T1_T2$Time,
                      MasterFrame_T1_T2$SampleID)
##models with no time
lme_no_time <- lapply(MasterFrame_T1_T2[,mets],
                      function(metInt, geno, group, ID){
                        ##full model
                        lme(metInt ~ geno*group, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GenoMDX, 
                      MasterFrame_T1_T2$GroupPred,
                      MasterFrame_T1_T2$SampleID)


```

Used saved models to perform association tests using ANOVA:

```{r}
##associations with genotype
p_geno <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_geno)
p_geno*length(all_mets)

##associations with group
p_group <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_group)
p_group*length(all_mets)

##associations with time
p_time <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_time)
p_time*length(all_mets)
```

## Correlation between these two metabolites

```{r}
cor(MasterFrameNoQC[,met_top], MasterFrameNoQC[,met_creatine])
cor(MasterFrame_T1_T2[,met_top], MasterFrame_T1_T2[,met_creatine])

cor(rank(MasterFrameNoQC[,met_top]), rank(MasterFrameNoQC[,met_creatine]))
cor(rank(MasterFrame_T1_T2[,met_top]), rank(MasterFrame_T1_T2[,met_creatine]))

```

## Test hypotheses relating to impact of genotype, group, and time for the top 50 metabolites from the T0 analysis

Fit models (include creatine and creatinine here again):

```{r}
top_peaks <- names(which(rank(sort(p_vals_T0))<=50))
top_peaks <- c(top_peaks, met_creatine, met_creatinine)

##full models
lme_full <- lapply(MasterFrame_T1_T2[,top_peaks],
                    function(metInt, geno, group, Times, ID){
                      ##full model
                      lme(metInt ~ geno*Times*group, random =~ 1 | ID,
                          method="ML");
                    }, 
                    MasterFrame_T1_T2$GenoMDX, MasterFrame_T1_T2$GroupPred, 
                    MasterFrame_T1_T2$Time,
                    MasterFrame_T1_T2$SampleID)
##models with no genotypes
lme_no_geno <- lapply(MasterFrame_T1_T2[,top_peaks],
                      function(metInt, group, Times, ID){
                        ##full model
                        lme(metInt ~ Times*group, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GroupPred, 
                      MasterFrame_T1_T2$Time,
                      MasterFrame_T1_T2$SampleID)
##models with no group
lme_no_group <- lapply(MasterFrame_T1_T2[,top_peaks],
                      function(metInt, geno, Times, ID){
                        ##full model
                        lme(metInt ~ Times*geno, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GenoMDX, 
                      MasterFrame_T1_T2$Time,
                      MasterFrame_T1_T2$SampleID)
##models with no time
lme_no_time <- lapply(MasterFrame_T1_T2[,top_peaks],
                      function(metInt, geno, group, ID){
                        ##full model
                        lme(metInt ~ geno*group, random =~ 1 | ID,
                            method="ML");
                      }, 
                      MasterFrame_T1_T2$GenoMDX, 
                      MasterFrame_T1_T2$GroupPred,
                      MasterFrame_T1_T2$SampleID)


```

Used saved models to perform association tests using ANOVA:

```{r}
length(all_mets)

##associations with genotype
p_geno <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_geno)
top_assoc_geno <- names(which(p_geno*length(all_mets) < 0.05))
length(top_assoc_geno)
p_geno[top_assoc_geno]
setdiff(top_peaks, top_assoc_geno)

##associations with group
p_group <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_group)
p_group <- sort(p_group)
##get the top metabolites in a vector
top_assoc_group <- names(which(sort(p_group)*length(all_mets) < 0.05))
length(top_assoc_group)
p_group[top_assoc_group]
##also look at other interesting metabolites here
p_group[met_top]
p_group[met_hypusine]

##associations with time
p_time <- mapply(function(mod1,mod2){anova(mod1,mod2)["p-value"][2,1]},
                 lme_full, lme_no_time)
##get the top metabolites in a vector
top_assoc_time <- names(which(sort(p_time)*length(all_mets) < 0.05))
length(top_assoc_time)
p_time[top_assoc_time]

##get which ones are associated with group but not time and vice-versa
setdiff(top_assoc_group, top_assoc_time)
setdiff(top_assoc_time, top_assoc_group)
##what about just among the top 10 associations with time?
setdiff(top_assoc_time[1:10], top_assoc_group)

```

Do a Venn diagram:

```{r venn_significant_overlaps, fig.height=5, fig.width=7}
all_peaks <- data.frame(matrix(0, nrow=length(top_peaks),
                               ncol=4))
colnames(all_peaks) <- c("Genotype (T0)",
                         "Treatment",
                         "Time",
                         "Genotype")
rownames(all_peaks) <- top_peaks
all_peaks[,1] <- 1
all_peaks[top_assoc_group,2] <- 1
all_peaks[top_assoc_time,3] <- 1
all_peaks[top_assoc_geno,4] <- 1

# v <- vennCounts(all_peaks)
# 
# vennDiagram(v)

v2 <- vennCounts(all_peaks[,2:4])

vennDiagram(v2,
            circle.col = c("orange",
                           "skyblue",
                           "pink1"))
```

## Now make some plots for the top metabolites associated with group

```{r, top_T0_geno_assoc_group_T1_T2}
MasterFrame_T0_WT <- filter(MasterFrame_T0, GenoMDX==0)
MasterFrame_T0_MDX <- filter(MasterFrame_T0, GenoMDX==1)

for(met in top_assoc_group)
{
  g1 <- ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met,"`",sep=""),
                                           group="SampleID", color="Group", linetype="Group")) +
    geom_point() +
    geom_line(aes(color=Genotype)) +
    scale_color_manual(name = "Genotype and group", 
                       breaks = c("group 1", "group 2", "WT", "mdx23"),
                       labels = c("group 1", "group 2", "WT", "mdx23"),
                       values = c(1, 3, 2, 4)) +
    ggtitle(paste("p_geno =", signif(p_geno[met],2),
                  ", p_time =", signif(p_time[met],2),
                  ", p_group =", signif(p_group[met],2)))
  
  g2 <- ggplot(MasterFrameNoQC, aes_string(x="Time", y=paste("`",met,"`",sep=""),
                                           group="SampleID", color="Group")) +
    geom_point() +
    geom_line(aes(color=Genotype)) +
    facet_grid(. ~ Group) +
    scale_color_manual(name = "Genotype and group", 
                       breaks = c("group 1", "group 2", "WT", "mdx23"),
                       labels = c("group 1", "group 2", "WT", "mdx23"),
                       values = c(1, 3, 2, 4)) +
    ggtitle(paste("p_geno =", signif(p_geno[met],2),
                  ", p_time =", signif(p_time[met],2),
                  ", p_group =", signif(p_group[met],2)))
  
  g3 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met,"`",sep=""),
                                             group="SampleID", color="Genotype", linetype="Group")) +
    geom_point() +
    geom_line() +
    scale_color_manual(name = "Genotype", 
                       breaks = c("WT", "mdx23"),
                       labels = c("WT", "mdx23"),
                       values = c(2, 4)) +
    ggtitle(paste("p_geno =", signif(p_geno[met],2),
                  ", p_time =", signif(p_time[met],2),
                  ", p_group =", signif(p_group[met],2)))
  
  ##better version of plot
  g3 <- ggplot(MasterFrame_T1_T2, aes_string(x="Time", y=paste("`",met,"`",sep=""),
                                             group="SampleID", color="Genotype", shape="Genotype", 
                                             linetype="Treatment")) +
    geom_point() +
    geom_line() +
    scale_color_manual(name = "Genotype", 
                       breaks = c("WT", "mdx"),
                       labels = c("WT", "mdx"),
                       values = c(2, 4)) +
    scale_shape_manual(name="Genotype",
                       breaks = c("WT", "mdx"),
                       labels = c("WT", "mdx"),
                       values = c(17, 16)) +
    ylab("Normalized, log2-transformed value") + 
    theme_bw() +
    theme(plot.title = element_text(hjust=0.5),
          legend.text = element_text(size=11))
  
  g4 <- ggplot(MasterFrame_T0_WT, aes_string(x="Group",y=paste("`",met,"`",sep=""))) +
    geom_boxplot() + geom_point()
  
  ##get p-value from t-test
  p4 <- t.test(MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 1",met],
               MasterFrame_T0_WT[MasterFrame_T0_WT$Group=="group 2",met])$p.value
  
  g4 <- g4 + ggtitle(paste("WT, p-value =", round(p4,3)))
  
  g5 <- ggplot(MasterFrame_T0_MDX, aes_string(x="Group",y=paste("`",met,"`",sep=""))) +
    geom_boxplot() + geom_point()
  
  ##get p-value from t-test
  p5 <- t.test(MasterFrame_T0_MDX[MasterFrame_T0_MDX$Group=="group 1",met],
               MasterFrame_T0_MDX[MasterFrame_T0_MDX$Group=="group 2",met])$p.value
  
  g5 <- g5 + ggtitle(paste("mdx, p-value =", round(p5,3)))

  print(g1)
  print(g2)
  print(g3)
  print(g4)
  print(g5)
}  

```

## Session info

```{r}
sessionInfo()
```
