---
title: "DMD mdx corticosteroids project: Merge statistical results for top peaks with HMDB annotations"
author: Simina Boca
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/")
```

## Read in files

```{r}
stat_res <- read.csv("../top_peaks_compare_genotypes_T0_urine.csv")
HMDB_pos <- read.csv("../top_peaks_compare_genotypes_T0_urine_HMDB_positive.csv")
HMDB_neg <- read.csv("../top_peaks_compare_genotypes_T0_urine_HMDB_negative.csv")
```

Look at the headers:

```{r}
head(stat_res)
head(HMDB_pos)
head(HMDB_neg)
```

## Process results so that they can be merged

Round the m/z value to 4 digits for all datasets:

```{r}
stat_res$mz <- round(stat_res$mz,4)
HMDB_pos$query_mass <- round(HMDB_pos$query_mass,4)
HMDB_neg$query_mass <- round(HMDB_neg$query_mass,4)
```

Rename columns with m/z value so they're the same for both datasets:

```{r}
colnames(stat_res)[colnames(stat_res) == "mz"] <- "peak_mz"
colnames(HMDB_pos)[colnames(HMDB_pos) == "query_mass"] <- 
  colnames(HMDB_neg)[colnames(HMDB_neg) == "query_mass"] <- 
  "peak_mz"
```

Add rank column to statistical results:

```{r}
##make sure they are sorted by increasing p-value
stat_res <- stat_res[order(stat_res$p_value),]
stat_res$rank <- 1:nrow(stat_res)

head(stat_res)
```

## Merge files

Merge all 3 files:

```{r}
HMDB_all <- rbind(HMDB_pos, HMDB_neg)
stat_res_annot <- merge(stat_res, HMDB_all,
                        all.x = TRUE)

head(stat_res_annot)
```

## Make some changes to output formatting and save it

Order output by rank:

```{r}
stat_res_annot <- stat_res_annot[order(stat_res_annot$rank),]

##make sure all ranks are included
identical(as.numeric(sort(unique(stat_res_annot$rank))),
          as.numeric(1:nrow(stat_res)))
```

Change some of the column names to match S4 File and save output:

```{r}
colnames(stat_res_annot) <- c("Peak m/z value",
                              "Peak mode",
                              "Peak retention time (seconds)",
                              "Log 2 fold change (mean difference)",
                              "DF",
                              "95% CI: lower bound", "95% CI: upper bound",
                              "T-value",
                              "P-value",
                              "Rank",
                              "Compound ID", "Compound name",
                              "KEGG ID", "Formula",
                              "Monoisotopic mass of putative metabolite",
                              "Putative adduct", "Putative adduct type",
                              "Adduct m/z value",
                              "Delta ppm")

##change some formatting, so it's the same as in the S4 file
stat_res_annot[,1] <- paste("m/z =", stat_res_annot[,1])
stat_res_annot[,2] <- paste("mode =", stat_res_annot[,2])
stat_res_annot[,3] <- paste("RT =", stat_res_annot[,3])

##check ppm range
range(stat_res_annot[,"Delta ppm"], na.rm = TRUE)

write.csv(x = stat_res_annot,
          row.names = FALSE,
          file = "../top_peaks_compare_genotypes_T0_urine_HMDB_both_modes.csv",
          quote = FALSE)
```

## Session info

```{r}
sessionInfo()
```
